---
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold"
fontsize: 14pt
classoption: "aspectratio=169"
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
## Render the pdf
##rmarkdown::render(input = "./11_1-Transforming_Variables.Rmd", output_file = "./11_1-Transforming_Variables.pdf")

##fontsizes for tables: ‘"tiny"’, ‘"scriptsize"’, ‘"footnotesize"’, ‘"small"’, ‘"normalsize"’, ‘"large"’, ‘"Large"’, ‘"LARGE"’, ‘"huge"’, ‘"Huge"’

library(tidyverse)
library(readxl)
library(stargazer)
##library(kableExtra)
library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)

#############################################################
## Stargazer code

## Regular Table
##stargazer(summary = FALSE, type = "latex", header = FALSE, rownames = FALSE, float = FALSE)

## Regression Results
##stargazer(res1, type = "latex", digits = 2, omit.stat = "rsq", header = FALSE, font.size = "footnotesize", star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, dep.var.caption = "", float = FALSE)

##covariate.labels = c("Mother's Education")
##dep.var.labels = "Earnings (2021)"

#############################################################
## Figures

## ```{r, fig.align = 'center', fig.asp=0.618, out.width = '100%', fig.width = 7}

##knitr::include_graphics()

#############################################################
## Side-by-side columns

## ::: columns
## :::: column
## Content in left column
## ::::
## :::: column
## Content in right column
## ::::
## :::

## Build bullets one-by-one
##\begin{enumerate}[<+->]

```


# Today's Agenda

\begin{large}

Extending the OLS Regression using Dataset 1

\begin{enumerate}

\item Dummy predictors
\item Categorical predictors
\item Transforming the variables
\item Transforming the model

\end{enumerate}

\end{large}

\vspace{.25in}

\begin{center}
Justin Leinaweaver (Spring 2022)
\end{center}



```{r, include = FALSE}
d <- read_excel("../../Data/Dataset-1/Dataset1-2020_Only.xlsx", na = "NA") %>%
  mutate(
    gdp_category = factor(gdp_category, levels = c("Under $100 billion", "Under $1 trillion", "Above $1 trillion")),
    pop_category = factor(pop_category, levels = c("Under 5 million", "Under 10 million", "Above 10 million")),
    manufacturing_category = factor(manufacturing_category, levels = c("Under 100k", "Under 300k", "Above 300k")),
    income_tax2 = if_else(income_tax == 1, "Income Tax", "No Income Tax"),
    log_gdp = log(gdp_millions*1e6),
    pop_millions = population_thousands / 1000,
    gdp_billions = gdp_millions/1000,
    log_pop = log(population_thousands * 1e3)
  )
```



# Transforming the Variables
::: columns
:::: column

\begin{center}
```{r, results = "asis"}
# Start by examining a bad model fit: GDP ~ bachelors
res1 <- lm(data = d, gdp_millions ~ bachelors)

stargazer(res1, digits = 2, dep.var.caption = "", star.cutoffs = .05, notes.append = FALSE, notes = "*p<0.05", omit.stat = "rsq", dep.var.labels = "GDP (millions)", covariate.labels = c("Bachelors (\\%)", "Intercept"), header = FALSE, float = FALSE, font.size = "footnotesize")
```
\end{center}

::::
:::: column
```{r, fig.align = 'center', fig.asp=0.85, out.width = '100%', fig.width = 5}
# Visualize the line fit plot (600x300)
d %>%
  ggplot(aes(x = bachelors, y = gdp_millions)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw() +
  labs(x = "Bachelors Degree (%)", y = "GDP (millions USD)")
```
::::
:::



# Transforming the Variables
::: columns
:::: column

\begin{center}
```{r, results = "asis"}
# Start by examining a bad model fit: GDP ~ bachelors
res2 <- lm(data = d, gdp_billions ~ bachelors)

stargazer(res2, digits = 2, dep.var.caption = "", star.cutoffs = .05, notes.append = FALSE, notes = "*p<0.05", omit.stat = "rsq", dep.var.labels = "GDP (billions)", covariate.labels = c("Bachelors (\\%)", "Intercept"), header = FALSE, float = FALSE, font.size = "footnotesize")
```
\end{center}

::::
:::: column
```{r, fig.align = 'center', fig.asp=0.85, out.width = '100%', fig.width = 5}
# Visualize the line fit plot (600x300)
d %>%
  ggplot(aes(x = bachelors, y = gdp_billions)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw() +
  labs(x = "Bachelors Degree (%)", y = "GDP (billions USD)")
```
::::
:::



# 
```{r, results = "asis"}
# Illustrate decimal transformations
d2 <- d %>%
  mutate(
    gdp_10billion = gdp_billions/10,
    gdp_100billion = gdp_billions/100
  )

##res1
##res2
res3 <- lm(data = d2, gdp_10billion ~ bachelors)
res4 <- lm(data = d2, gdp_100billion ~ bachelors)

stargazer(res1, res2, res3, res4, digits = 2, dep.var.caption = "", star.cutoffs = .05, notes.append = FALSE, notes = "*p<0.05", omit.stat = "rsq", dep.var.labels = c("GDP (millions)", "(billions)", "(10 billion)", "(100 billion)"), covariate.labels = c("Bachelors (\\%)", "Intercept"), header = FALSE, float = FALSE, font.size = "footnotesize")
```


# Transformation 1: Shift the Decimal Point
```{r, fig.align = 'center', fig.asp=0.618, out.height = '80%', fig.width = 5.5}
# Visualize the residual plot (600x300)
d %>%
  add_residuals(res2) %>%
  add_predictions(res2) %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  theme_bw() +
  labs(x = "Predicted GDP (Billions USD)", y = "Residuals") +
  coord_cartesian(ylim = c(-1000, 3000)) +
  geom_hline(yintercept = 0, color = "red")
```



# Transformation 1: Shift the Decimal Point
::: columns
:::: column
```{r, fig.align = 'center', fig.asp=0.85, out.height = '80%', fig.width = 5.5}
# Visualize the residual plot (600x300)
resid1 <- d %>%
  add_residuals(res2) %>%
    add_predictions(res2)

resid1 %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  theme_bw() +
  labs(x = "Predicted GDP (Billions USD)", y = "Residuals") +
  coord_cartesian(ylim = c(-1000, 3000)) +
  geom_hline(yintercept = 0, color = "red")
```
::::
:::: column

\vspace{1in}

\begin{Large}
\begin{center}

`r sum(resid1$resid < 0)` / 50 (`r (sum(resid1$resid < 0)/50)*100` \%) are below the zero line.

\vspace{.2in}

Aim is 50\%-50\%

\end{center}
\end{Large}
::::
:::



# Transformation 2: Natural Logarithms

\begin{Large}

\begin{enumerate}

\item Create a new variable: "log GDP"

\begin{itemize}

\begin{large}

\item Excel function for the natural log is = LN()

\item GDP in Dataset 1 is in millions so you have to convert it to dollars first!

\item = LN( GDP * 1e6 )

\end{large}

\end{itemize}

\item Practice: Regress log GDP on bachelors

\end{enumerate}

\end{Large}



# 
::: columns
:::: column
```{r, results = "asis"}
##### Bachelors #####
res1 <- d %>% lm(data = ., gdp_billions ~ bachelors)
res2 <- d %>% lm(data = ., log_gdp ~ bachelors)

## Save coefficients
x1 <- round(coef(res1), 2)
x2 <- round(coef(res2), 2)

stargazer(res1, res2, digits = 2, dep.var.caption = "", star.cutoffs = .05, notes.append = FALSE, notes = "*p<0.05", omit.stat = "rsq", dep.var.labels = c("(billions)", "(log)"), covariate.labels = c("Bachelors (\\%)", "Intercept"), header = FALSE, float = FALSE, font.size = "footnotesize")
```
::::
:::: column

\vspace{.4in}

```{r, fig.align = 'center', fig.asp=0.8, out.height = '65%', fig.width = 5}
# Visualize the line fit plot (600x300)
d %>%
  ggplot(aes(x = bachelors, y = log_gdp)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw() +
  labs(x = "Bachelors Degree (%)", y = "log (GDP)")
```
::::
:::



# 
::: columns
:::: column
```{r, fig.align = 'center', fig.asp=0.8, out.height = '70%', fig.width = 5}
# Visualize the residual plot (600x300)
resid11 <- d %>%
  add_residuals(res1) %>%
    add_predictions(res1)

resid11 %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  theme_bw() +
  labs(x = "Predicted GDP (billions USD)", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")
```

\begin{center}

\begin{large}

Above the line = `r sum(resid11$resid > 0)` (`r round(sum(resid11$resid > 0)/50*100,0)`\%)

Below the line = `r sum(resid11$resid < 0)` (`r round(sum(resid11$resid < 0)/50*100,0)`\%)

\end{large}

\end{center}

::::
:::: column

```{r, fig.align = 'center', fig.asp=0.8, out.height = '70%', fig.width = 5}
# Visualize the residual plot (600x300)
resid12 <- d %>%
  add_residuals(res2) %>%
    add_predictions(res2)

resid12 %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  theme_bw() +
  labs(x = "Predicted GDP (log)", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")
```

\begin{center}
\begin{large}

Above the line = `r sum(resid12$resid > 0)` (`r round(sum(resid12$resid > 0)/50*100,0)`\%)

Below the line = `r sum(resid12$resid < 0)` (`r round(sum(resid12$resid < 0)/50*100,0)`\%)

\end{large}
\end{center}
::::
:::




# Transformation 2: Natural Logarithms

```{r, fig.align = 'center', fig.asp=0.5, out.height = '90%', fig.width = 7}
## Plot GDP vs Log GDP
d %>%
  pivot_longer(cols = c(gdp_billions, log_gdp), names_to = "Version", values_to = "Values") %>%
  mutate(
    Version = if_else(Version == "gdp_billions", "GDP (Billions USD)", "log of GDP (USD)")
  ) %>%
  ggplot(aes(x = Values)) +
  geom_histogram(bins = 15, fill = "darkgrey", color = "white") +
  theme_bw() +
  labs(x = "", y = "") +
  facet_wrap(~ Version, scales = "free", ncol = 2)
```



# Transformation 2: Natural Logarithms

```{r, fig.align = 'center', fig.asp=0.5, out.height = '90%', fig.width = 7}
## Plot GDP vs Log GDP (add mean line)
new1 <- tibble(
  Version = c("GDP (Billions USD)", "log of GDP (USD)"),
  Values = c(mean(d$gdp_billions), mean(d$log_gdp))
)

d %>%
  pivot_longer(cols = c(gdp_billions, log_gdp), names_to = "Version", values_to = "Values") %>%
  mutate(
    Version = if_else(Version == "gdp_billions", "GDP (Billions USD)", "log of GDP (USD)"),
    Mean = mean(Values)
  ) %>% 
  ggplot(aes(x = Values)) +
  geom_histogram(bins = 15, fill = "darkgrey", color = "white") +
  geom_vline(data = new1, aes(xintercept = Values), color = "red", size = 3) +
  theme_bw() +
  labs(x = "", y = "") +
  facet_wrap(~ Version, scales = "free", ncol = 2)
```



# Transformation 2: Natural Logarithms

```{r, fig.align = 'center', fig.asp=0.5, out.height = '90%', fig.width = 7}
## Plot GDP vs Log GDP, add highlights to show how log stretches out differences at the low end
d %>%
  mutate(
    color_in = if_else(gdp_billions < 110, "1", "0")
  ) %>%
  pivot_longer(cols = c(gdp_billions, log_gdp), names_to = "Version", values_to = "Values") %>%
  mutate(
    Version = if_else(Version == "gdp_billions", "GDP (Billions USD)", "log of GDP (USD)")
  ) %>%
  ggplot(aes(x = Values, fill = color_in)) +
  geom_histogram(bins = 15, color = "white") +
  theme_bw() +
  labs(x = "", y = "") +
  facet_wrap(~ Version, scales = "free", ncol = 2) +
  guides(fill = FALSE) +
  scale_fill_manual(values = c("darkgrey", "orange"))
```



# Transformation 2: Natural Logarithms

```{r, fig.align = 'center', fig.asp=0.5, out.height = '90%', fig.width = 7}
## Plot GDP vs Log GDP, add highlights to show how log condenses differences at the high end
d %>%
  mutate(
    color_in = if_else(gdp_billions > 1500, "1", "0")
  ) %>%
  pivot_longer(cols = c(gdp_billions, log_gdp), names_to = "Version", values_to = "Values") %>%
  mutate(
    Version = if_else(Version == "gdp_billions", "GDP (Billions USD)", "log of GDP (USD)")
  ) %>%
  ggplot(aes(x = Values, fill = color_in)) +
  geom_histogram(bins = 15, color = "white") +
  theme_bw() +
  labs(x = "", y = "") +
  facet_wrap(~ Version, scales = "free", ncol = 2) +
  guides(fill = FALSE) +
  scale_fill_manual(values = c("darkgrey", "red"))
```



# Transformation 2: Natural Logarithms

```{r, fig.align = 'center', fig.asp=0.5, out.height = '85%', fig.width = 6}
## Visualize the natural log scale
tibble(
  x = 1:20, #1e3,
  log_x = log(x)
) %>%
  ggplot(aes(x = x, y = log_x)) +
  geom_line() +
  theme_minimal() +
  labs(x = "X", y = "Natural Log of X") #+
  scale_x_continuous(breaks = 1:20)
```



# Transformation 2: Natural Logarithms

```{r, fig.align = 'center', fig.asp=0.5, out.height = '85%', fig.width = 6}
## Visualize the natural log scale
tibble(
  x = 1:1e4,
  log_x = log(x)
) %>%
  ggplot(aes(x = x, y = log_x)) +
  geom_line() +
  theme_minimal() +
  labs(x = "X", y = "Natural Log of X")
```



# Transformation 2: Natural Logarithms

\begin{LARGE}

The natural log scale = multiplying by e

\begin{itemize}

\item $e$ is Euler's Number (2.718282...)

\item Typically written as $log_e$ X or ln X

\end{itemize}

\vspace{.25in}

Transform back to linear scale using eX

\end{LARGE}



# Transformation 2: Natural Logarithms

\begin{LARGE}

GDP for State X = \$5,000 or 8.5 (ln)

\begin{itemize}

\item $log_e$ 5,000 $\approx$ 8.5

\item $e^{8.5} \approx$ 5,000 

\item Excel function is = EXP()

\end{itemize}

\end{LARGE}



# 
```{r, fig.align = 'center', fig.asp=0.618, out.height = '70%', fig.width = 5}
p1 <- d %>%
  ggplot(aes(x = log_gdp)) +
  geom_histogram(bins = 15, fill = "darkgrey", color = "white") +
  theme_minimal() +
  labs(x = "", y = "", title = "Log (GDP)")

p1 +
    geom_vline(xintercept = c(24, 29), color = "red", size = 1.3)
```

\begin{large}
\begin{itemize}

\item $e^{24}$ is approximately \$26 billion

\item $e^{29}$ is approximately \$3.9 trillion

\end{itemize}
\end{large}



# 
::: columns
:::: column
```{r, fig.align = 'center', fig.asp=0.618, out.height = '60%', fig.width = 5}
p1 +
    geom_vline(xintercept = c(24, 25), color = "red", size = 1.3)
```

\begin{center}
\begin{itemize}

\item $e^{24}$ $\approx$ \$26 billion

\item $e^{25}$ $\approx$ \$72 billion

\end{itemize}
\end{center}

::::
:::: column

\vspace{.5in}

\begin{center}
\begin{large}
One unit on the LN scale: 

\begin{itemize}
\item value x $e$
\end{itemize}

\vspace{.5in}

General rule of thumb:

\begin{itemize}
\item value x 3
\end{itemize}

\end{large}
\end{center}
::::
:::



#
::: columns
:::: column

\vspace{.5in}

\begin{center}
\begin{LARGE}
\textbf{Making Point Estimates}
\end{LARGE}
\end{center}
::::
:::: column

```{r, results = "asis"}
stargazer(res2, digits = 2, dep.var.caption = "", star.cutoffs = .06, notes.append = FALSE, notes = "*p<0.06", omit.stat = "rsq", dep.var.labels = c("(log GDP)"), covariate.labels = c("Bachelors (\\%)"), header = FALSE, float = FALSE, font.size = "footnotesize")
```
::::
:::

\vspace{.15in}

\begin{center}
\textbf{ln (Outcome) = Intercept + Coefficient * (Predictor)}
\end{center}



#
::: columns
:::: column

\vspace{.5in}

\begin{center}
\begin{LARGE}
\textbf{Making Point Estimates}
\end{LARGE}
\end{center}
::::
:::: column

```{r, results = "asis"}
stargazer(res2, digits = 2, dep.var.caption = "", star.cutoffs = .06, notes.append = FALSE, notes = "*p<0.06", omit.stat = "rsq", dep.var.labels = c("(log GDP)"), covariate.labels = c("Bachelors (\\%)"), header = FALSE, float = FALSE, font.size = "footnotesize")
```
::::
:::

\vspace{.15in}

\begin{center}
\textbf{ln (GDP) = `r x2[1]` + `r x2[2]` * (Bachelors)}
\end{center}



#
::: columns
:::: column

\vspace{.5in}

\begin{center}
\begin{LARGE}
\textbf{Making Point Estimates}
\end{LARGE}
\end{center}
::::
:::: column

```{r, results = "asis"}
stargazer(res2, digits = 2, dep.var.caption = "", star.cutoffs = .06, notes.append = FALSE, notes = "*p<0.06", omit.stat = "rsq", dep.var.labels = c("(log GDP)"), covariate.labels = c("Bachelors (\\%)"), header = FALSE, float = FALSE, font.size = "footnotesize")
```
::::
:::

\vspace{.15in}

\begin{center}
\textbf{ln (GDP) = `r x2[1]` + `r x2[2]` * `r round(mean(d$bachelors),2)` = `r round(x2[1]+x2[2]*mean(d$bachelors), 2)`}
\end{center}




#
::: columns
:::: column

\vspace{.5in}

\begin{center}
\begin{LARGE}
\textbf{Making Point Estimates}
\end{LARGE}
\end{center}
::::
:::: column

```{r, results = "asis"}
stargazer(res2, digits = 2, dep.var.caption = "", star.cutoffs = .06, notes.append = FALSE, notes = "*p<0.06", omit.stat = "rsq", dep.var.labels = c("(log GDP)"), covariate.labels = c("Bachelors (\\%)"), header = FALSE, float = FALSE, font.size = "footnotesize")

pred1 <- as.numeric(round(x2[1]+x2[2]*mean(d$bachelors), 2))
```
::::
:::

\vspace{.15in}

\begin{center}
\textbf{GDP = `r pred1` (ln) = $e^{`r pred1`}$ = `r round(exp(pred1)/1e9,1)` Billion USD}
\end{center}



# Making Point Estimates

\begin{center}
\begin{Large}

ln(Outcome) = Intercept + Coefficient * (Predictor)

\vspace{.5in}

Outcome = $e^{Intercept + Coefficient * (Predictor)}$

\end{Large}
\end{center}



# Practice

\begin{Large}

\textbf{Dataset 2: Do wealthier countries live longer?}

\begin{itemize}
\item Regress life expectancy on GDP
\item Regress life expectancy on log(GDP)
\end{itemize}

\end{Large}


#
::: columns
:::: column

\vspace{.2in}

```{r, results = "asis"}
d <- read_excel("../../Data/Dataset-2/Dataset2-2019_only.xlsx") %>%
    mutate(
        log_gdp = log(gdp_current)
    )

res5 <- lm(data = d, life_expectancy ~ gdp_current)
res6 <- lm(data = d, life_expectancy ~ log_gdp)

stargazer(res5, res6, digits = 2, dep.var.caption = "", star.cutoffs = .05, notes.append = FALSE, notes = "*p<0.05", omit.stat = "rsq", dep.var.labels = "Life Expectancy", covariate.labels = c("GDP", "log(GDP)"), header = FALSE, float = FALSE, font.size = "scriptsize")
```
::::
:::: column
```{r, fig.align = 'center', fig.asp=0.618, out.height = '45%', fig.width = 5}
d %>%
    ggplot(aes(x = gdp_current, y = life_expectancy)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    theme_bw() +
    labs(x = "GDP (current USD)", y = "Life Expectancy")

d %>%
    ggplot(aes(x = log_gdp, y = life_expectancy)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    theme_bw() +
    labs(x = "log (GDP)", y = "Life Expectancy")
```
::::
:::


