---
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold"
fontsize: 14pt
classoption: "aspectratio=169"
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
- \setbeamertemplate{page number in head/foot}[]{}
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
## Render the pdf
##rmarkdown::render(input = "./14_15-Analyzing_Sales_Data.Rmd", output_file = "./14_15-Analyzing_Sales_Data.pdf")

##fontsizes for tables: ‘"tiny"’, ‘"scriptsize"’, ‘"footnotesize"’, ‘"small"’, ‘"normalsize"’, ‘"large"’, ‘"Large"’, ‘"LARGE"’, ‘"huge"’, ‘"Huge"’

library(tidyverse)
library(readxl)
library(stargazer)
##library(kableExtra)
library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)

#############################################################
## Stargazer code

## Regular Table
##stargazer(summary = FALSE, type = "latex", header = FALSE, rownames = FALSE, float = FALSE)

## Regression Results
##stargazer(res1, type = "latex", digits = 2, omit.stat = "rsq", header = FALSE, font.size = "footnotesize", star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, dep.var.caption = "", float = FALSE)

##covariate.labels = c("Mother's Education")
##dep.var.labels = "Earnings (2021)"

#############################################################
## Figures

## ```{r, fig.align = 'center', fig.asp=0.618, out.height = '90%', fig.width = 5}

##knitr::include_graphics()

#############################################################
## Side-by-side columns

## ::: columns
## :::: column
## Content in left column
## ::::
## :::: column
## Content in right column
## ::::
## :::

## Build bullets one-by-one
##\begin{enumerate}[<+->]

```


# Today's Agenda

\begin{LARGE}

Weeks 14 and 15: The Final Project

\begin{itemize}
\item Data is now available on Moodle
\end{itemize}
\end{LARGE}

\vspace{.5in}

\begin{center}
Justin Leinaweaver (Spring 2022)
\end{center}



# The Final Report
\begin{center}
```{r, results = "asis"}
d <- read_excel("../../Data/00-Final_Project/Final_Project_Data_SP22.xlsx") %>%
    mutate(
        Male = if_else(Gender == "Male", 1, 0),
        Spring = if_else(Quarter == "Spring", 1, 0),
        Summer = if_else(Quarter == "Summer", 1, 0),
        Fall = if_else(Quarter == "Fall", 1, 0),
        Winter = if_else(Quarter == "Winter", 1, 0),
        North = if_else(Region == "North", 1, 0),
        South = if_else(Region == "South", 1, 0),
        East = if_else(Region == "East", 1, 0),
        West = if_else(Region == "West", 1, 0)
    )

d %>%
    select(Salesperson:Gender) %>%
    filter(Year == 2016, Time == 1) %>%
    slice(c(seq(1,20,3))) %>%
    mutate_if(is.numeric, round, 2) %>%
    arrange(Year, Time) %>%
    stargazer(summary = FALSE, header = FALSE, float = FALSE, rownames = FALSE, font.size = "small")    
```
\end{center}

\vspace{.2in}

\begin{center}
\begin{Large}
Quarterly sales data for 16 salespeople at our hypothetical company
\end{Large}
\end{center}



# The Final Report
\begin{center}
```{r, results = "asis"}
d %>%
    select(Salesperson:Region) %>%
    slice(1:7) %>%
    mutate_if(is.numeric, round, 2) %>%
    arrange(Year, Time) %>%
    stargazer(summary = FALSE, header = FALSE, float = FALSE, rownames = FALSE, font.size = "small")
```
\end{center}

\begin{center}
\begin{Large}
Each salesperson has 20 observations

(Sales across 4 quarters x 5 years)
\end{Large}
\end{center}



# The Final Report
\begin{center}
```{r, results = "asis"}
d %>%
    select(Salesperson, Gender:Age) %>%
    distinct %>%
    slice(1:7) %>%
    mutate_if(is.numeric, round, 2) %>%
    stargazer(summary = FALSE, header = FALSE, float = FALSE, rownames = FALSE, font.size = "small")
```

\begin{Large}
Dataset provides basic demographic info for each salesperson
\end{Large}
\end{center}


# The Final Report: Three Sections

\begin{LARGE}

\begin{itemize}

\item Each section requires an argument
\begin{itemize}
\begin{large}
\item Introduction, Body and Conclusion
\end{large}
\end{itemize}

\item Every claim requires evidence
\begin{itemize}
\begin{large}
\item Descriptive statistics, visualizations, tables, etc.
\end{large}
\end{itemize}

\item Assignment details on Moodle
\end{itemize}
\end{LARGE}
