---
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold"
fontsize: 14pt
classoption: "aspectratio=169"
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
- \setbeamertemplate{page number in head/foot}[]{}
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
## Render the pdf
##rmarkdown::render(input = "./14_15-Analyzing_Sales_Data.Rmd", output_file = "./14_15-Analyzing_Sales_Data.pdf")

##fontsizes for tables: ‘"tiny"’, ‘"scriptsize"’, ‘"footnotesize"’, ‘"small"’, ‘"normalsize"’, ‘"large"’, ‘"Large"’, ‘"LARGE"’, ‘"huge"’, ‘"Huge"’

library(tidyverse)
library(readxl)
library(stargazer)
##library(kableExtra)
library(modelr)

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)

#############################################################
## Stargazer code

## Regular Table
##stargazer(summary = FALSE, type = "latex", header = FALSE, rownames = FALSE, float = FALSE)

## Regression Results
##stargazer(res1, type = "latex", digits = 2, omit.stat = "rsq", header = FALSE, font.size = "footnotesize", star.cutoffs = .05, notes = "*p < 0.05", notes.append = FALSE, dep.var.caption = "", float = FALSE)

##covariate.labels = c("Mother's Education")
##dep.var.labels = "Earnings (2021)"

#############################################################
## Figures

## ```{r, fig.align = 'center', fig.asp=0.618, out.height = '90%', fig.width = 5}

##knitr::include_graphics()

#############################################################
## Side-by-side columns

## ::: columns
## :::: column
## Content in left column
## ::::
## :::: column
## Content in right column
## ::::
## :::

## Build bullets one-by-one
##\begin{enumerate}[<+->]

```


# Today's Agenda

\begin{Large}

Weeks 14 and 15: The Final Project

\begin{itemize}
\item Assignment (w/ data) available on Moodle
\end{itemize}
\end{Large}

\vspace{.5in}

\begin{center}
Justin Leinaweaver (Spring 2022)
\end{center}



# Final Report: The Data
\begin{center}
\begin{Large}
Quarterly sales data for 16 salespeople at our hypothetical company
\end{Large}
\end{center}

```{r}
d <- read_excel("../../Data/00-Final_Project/Final_Project_Data_SP22.xlsx") %>%
    mutate(
        Male = if_else(Gender == "Male", 1, 0),
        Spring = if_else(Quarter == "Spring", 1, 0),
        Summer = if_else(Quarter == "Summer", 1, 0),
        Fall = if_else(Quarter == "Fall", 1, 0),
        Winter = if_else(Quarter == "Winter", 1, 0),
        North = if_else(Region == "North", 1, 0),
        South = if_else(Region == "South", 1, 0),
        East = if_else(Region == "East", 1, 0),
        West = if_else(Region == "West", 1, 0)
    )
```

\vspace{.2in}

**Sales Team:** `r sort(unique(d$Salesperson))`

\vspace{.2in}

**Time Period:** Quarterly data from `r head(unique(d$Year), n = 1)` to `r tail(unique(d$Year), n = 1)` (Winter, Spring, Summer and Fall)





# Final Report: The Data
\begin{center}
```{r, results = "asis"}
d %>%
    select(Salesperson:Sales) %>%
    slice(1:7) %>%
    mutate_if(is.numeric, round, 2) %>%
    arrange(Year, Time) %>%
    stargazer(summary = FALSE, header = FALSE, float = FALSE, rownames = FALSE, font.size = "small")
```
\end{center}

\begin{center}
\begin{Large}
Each salesperson has 20 observations

(Sales across 4 quarters x 5 years)
\end{Large}
\end{center}



# Final Report: The Data
\begin{center}
```{r, results = "asis"}
d |>
  select(Salesperson, Gender:Age) |>
  slice(seq(1, 320, 21)) |>
  arrange(Salesperson) |>
  slice(1:7) |>
  stargazer(summary = FALSE, header = FALSE, float = FALSE, rownames = FALSE, font.size = "small")
```

\begin{Large}
Dataset provides basic demographic info for each salesperson
\end{Large}
\end{center}




# Final Report: The Data
\begin{center}
```{r, results = "asis"}
d %>%
    select(Salesperson, Year, Gender:Age) %>%
    distinct %>%
    slice(1:7) %>%
    mutate_if(is.numeric, round, 2) %>%
    stargazer(summary = FALSE, header = FALSE, float = FALSE, rownames = FALSE, font.size = "small")
```

\begin{Large}
Note that age increases across time!
\end{Large}
\end{center}



# Final Report: Three Sections
\Large

1. Each section requires a complete argument
    - \large Introduction, body and conclusion
    - Every claim requires statistical evidence

2. Assignment details on Moodle

3. Almost 6 hours of in-class time to work
    - \large Wk 14, 15 and the final exam period



# Final Report: Section 1
\large

**1. How has our sales staff performed over the last year (2021)? Who are our strongest salespeople? Who are our weakest?**

- Consider performance **both** in terms of **total sales** and in terms of **year-to-year growth**.

- Make an argument and support it with visualizations and descriptive statistics.




# Final Report: Section 2

**What is the profile of our most successful salespeople? Build a multiple OLS regression in order to identify the strongest predictors that explain quarterly sales.**

- Consider each of the following predictors for your "best" model: Age, education, gender, region of the country, time and the quarter of the year.

- Be sure to include a properly formatted regression table, evaluate your model with all five textbook steps (including a residual plot) and make a marginal effects plot to visualize the strongest predictor of sales.




# Final Report: Section 3

**3. What is your best forecast of our company-wide sales for the next quarter (Winter 2022)? How confident are you in this projection?**

- You will need to combine the quarterly sales figures for all of our salespeople to calculate a quarterly total for the business as a whole.

- You must consider the following methods: 1. naive, 2. MA-3, and 3. a linear prediction model (e.g. quadratic, cubic and season dummies)

- Be sure to provide your forecast results in both a table and a visualization.